WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2015-11-19 ms.tex
Fix incorrect figure reference
Switch to absolute magnitude

The \acl{LIGO} (\acsu{LIGO}; \cite{AdvancedLIGOHarry,AdvancedLIGO}) and Virgo \cite{AdvancedVirgo} have just begun taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations. They will ultimately increase their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010. They form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India \cite{LIGOIndia}.

The most readily detectable sources of \acp{GW} include binary neutron star mergers, with 0.4\nobreakdashes--400 events per year within the reach of Advanced LIGO at final design sensitivity~\cite{rates}. These binary systems are not only efficient sources of \acp{GW}, but also potential sources of detectable \ac{EM} transients from the aftermath of the tidal disruption of the \acp{NS}. \citet{MostPromisingEMCounterpart} argue that the most promising \ac{EM} counterparts are the hypothesized optical/infrared ``kilonovae'' powered by the radioactivate decay of $r$\nobreakdashes-process elements synthesized within the neutron\nobreakdashes-rich ejecta. These are expected to be faint, red, and peak rapidly, reaching only $R \sim 24$~mag within a week, though rising several magnitudes brighter in the infrared \citep{KilonovaHighOpacities}.

@@

The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and electromagnetic emission from compact binaries. This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}, \aclu{MBTA}; \cite{LIGOVirgoInspiralPipelineComparison}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities. Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.

@@

The resulting map is a tree structure that describes a mesh of pixels with different resolutions. An example \ac{BAYESTAR} subdivision is shown in Figure~\ref{fig:adaptive-illustration}. To convert this mesh into a FITS image, 
we traverse the tree and flatten it into the highest resolution represented. The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$~deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values. For the purpose of delivery to observers, therefore, the output is always \texttt{gzip}\nobreakdashes-compressed with a ratio of $\approx 250:1$.}

>>>>

The \acl{LIGO} (\acsu{LIGO}; \cite{AdvancedLIGOHarry,AdvancedLIGO}) and Virgo \cite{AdvancedVirgo} have just begun taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations. They will ultimately increase their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010. They form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India~\cite{LIGOIndia}.

The most readily detectable sources of \acp{GW} include binary neutron star mergers, with 0.4\nobreakdashes--400 events per year within the reach of Advanced LIGO at final design sensitivity~\cite{rates}. These binary systems are not only efficient sources of \acp{GW}, but also potential sources of detectable \ac{EM} transients from the aftermath of the tidal disruption of the \acp{NS}. \citet{MostPromisingEMCounterpart} argue that the most promising \ac{EM} counterparts are the hypothesized optical/infrared ``kilonovae'' powered by the radioactivate decay of $r$\nobreakdashes-process elements synthesized within the neutron\nobreakdashes-rich ejecta. These are expected to be faint, red, and peak rapidly, reaching an absolute magnitude of only $M_R \sim -13$ within a week, though rising several magnitudes brighter in the infrared~\citep{KilonovaHighOpacities}.

@@

The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and electromagnetic emission from compact binaries. This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}, \aclu{MBTA}; \cite{LIGOVirgoInspiralPipelineComparison}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities. Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run~\cite{SiderySkyLocalizationComparison}.

@@

The resulting map is a tree structure that describes a mesh of pixels with different resolutions. An example \ac{BAYESTAR} subdivision is shown in Figure~\ref{fig:adaptive-mesh}. To convert this mesh into a FITS image, we traverse the tree and flatten it into the highest resolution represented. The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$~deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values. For the purpose of delivery to observers, therefore, the output is always \texttt{gzip}\nobreakdashes-compressed with a ratio of $\approx 250:1$.}
